---
title: "RISG_CADO_Juv_Outliers"
author: "Amy Zyck"
date: '2024-11-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Fst and PCA of candidate juvenile loci 

## Organizing vcf files

In terminal:

In `juv_ddocent` directory, make a new directory for Pop and Sea Gen analyses

```
$ mkdir Juv_Selection_Analysis 
$ cd Juv_Selection_Analysis
```

Make a sub directory for the outlier SNP sets and link vcf files 

```
$ mkdir Juv_Outlier
$ cd Juv_Outlier

$ ln -s ../../Juv_OutlierDetection/outlierloci_juv.recode.vcf .
$ ln -s ../../Juv_OutlierDetection/outlierloci_juv_db.recode.vcf .
$ ln -s ../../Juv_OutlierDetection/outlierloci_juv_lfmm.recode.vcf .
```

```{r}
#Loading all the necessary packages
library(adegenet)
library(vcfR)
library("radiator") # Conversion from vcf to a lot of other formats
library("dplyr")
library("hierfstat")
library("ggplot2") #For plotting
library("reshape2") #For plotting
#library("plyr")
library("cowplot") #For plotting manuscript figs
library(PCAviz)  #Visualizing output of PCA
#library("stringr")
library(tidyverse)
library("poppr")
library(eulerr)
```


## Making files

```{r}
# Setting up strata file - population info and environmental data 
strata_juv <- read.table("../strata_cado", header=TRUE)

# Making new version of strata file with one row per population
strata_juv_pops <- strata_juv %>% distinct(Population, .keep_all = TRUE)
```

```{r}
#Providing population names for plotting
pop_order_juv <- c("Con1","Con2","Con3","Hi1","Hi2","Hi3")
```

All outliers

**Make genind object**

```{r}
# VCF file with juv SNP dataset
my_vcf_out_juv <- read.vcfR("outlierloci_juv.recode.vcf")
```

```{r}
rad_out_juv.filt <- vcfR2genind(my_vcf_out_juv, strata = strata_juv, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))

rad_out_juv.filt
```

**Make hierfstat object**

```{r}
hf_out_juv.filt <- genind2hierfstat(rad_out_juv.filt, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))
```

## Pairwise Fst  

```{r}
fst_out_juv.mat <- pairwise.WCfst(hf_out_juv.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_juv_fst <- boot.ppfst(hf_out_juv.filt,nboot = 1000)
boot5_out_juv_fst <- boot.ppfst(hf_out_juv.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_juv_ll <- boot5_out_juv_fst$ll
```

```{r}
gindF_out_juv.fst.mat.triN <- as.matrix(fst_out_juv.mat)
colnames(gindF_out_juv.fst.mat.triN) <- pop_order_juv
rownames(gindF_out_juv.fst.mat.triN) <- pop_order_juv
```

```{r}
meltedN_out_juv <- melt(gindF_out_juv.fst.mat.triN, na.rm =TRUE)
round(gindF_out_juv.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_juv$value)
```

```{r}
#Plotting Pairwise fst
out_juv <- ggplot(data = meltedN_out_juv, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) juv outlier SNPs", atop(italic("N = 120, L = 346"), ""))))+
  labs( x = "Treatment", y = "Treatment") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_juv
```

## PCA 

```{r}
library("PCAviz")  #Visualizing output of PCA
library("cowplot") #Used with PCAviz
```

```{r}
# Create color scheme
cols_juv = c("blue1","lightblue","darkblue","red","red3","darkred")
```

```{r}
# Plot a PCA for the all outlier loci
x_out_juv <- scaleGen(rad_out_juv.filt, NA.method = "mean")
out_juv_pca <- dudi.pca(x_out_juv, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_juv <- round((out_juv_pca$eig / (sum(out_juv_pca$eig))) * 100, 2)

pca_dat_out_juv <- as.data.frame(out_juv_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_juv.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Con1", "Con2", "Con3", "Hi1", "Hi2", "Hi3")))


ggplot(data = pca_dat_out_juv) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_juv[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_juv[2], "%")) +
    coord_fixed(ratio = 1)
```


LFMM outliers

**Make genind object**

```{r}
# VCF file with juv_lfmm SNP dataset
my_vcf_out_juv_lfmm <- read.vcfR("outlierloci_juv_lfmm.recode.vcf")
```

```{r}
rad_out_juv_lfmm.filt <- vcfR2genind(my_vcf_out_juv_lfmm, strata = strata_juv, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))

rad_out_juv_lfmm.filt
```

**Make hierfstat object**

```{r}
hf_out_juv_lfmm.filt <- genind2hierfstat(rad_out_juv_lfmm.filt, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))
```

## Pairwise Fst  

```{r}
fst_out_juv_lfmm.mat <- pairwise.WCfst(hf_out_juv_lfmm.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_juv_lfmm_fst <- boot.ppfst(hf_out_juv_lfmm.filt,nboot = 1000)
boot5_out_juv_lfmm_fst <- boot.ppfst(hf_out_juv_lfmm.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_juv_lfmm_ll <- boot5_out_juv_lfmm_fst$ll
```

```{r}
gindF_out_juv_lfmm.fst.mat.triN <- as.matrix(fst_out_juv_lfmm.mat)
colnames(gindF_out_juv_lfmm.fst.mat.triN) <- pop_order_juv
rownames(gindF_out_juv_lfmm.fst.mat.triN) <- pop_order_juv
```

```{r}
meltedN_out_juv_lfmm <- melt(gindF_out_juv_lfmm.fst.mat.triN, na.rm =TRUE)
round(gindF_out_juv_lfmm.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_juv_lfmm$value)
```

```{r}
#Plotting Pairwise fst
out_juv_lfmm <- ggplot(data = meltedN_out_juv_lfmm, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) juv_lfmm outlier SNPs", atop(italic("N = 120, L = 24"), ""))))+
  labs( x = "Treatment", y = "Treatment") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_juv_lfmm
```

## PCA 

```{r}
library("PCAviz")  #Visualizing output of PCA
library("cowplot") #Used with PCAviz
```

```{r}
# Create color scheme
cols_juv_lfmm = c("blue1","lightblue","darkblue","red","red3","darkred")
```

```{r}
# Plot a PCA for the all outlier loci
x_out_juv_lfmm <- scaleGen(rad_out_juv_lfmm.filt, NA.method = "mean")
out_juv_lfmm_pca <- dudi.pca(x_out_juv_lfmm, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_juv_lfmm <- round((out_juv_lfmm_pca$eig / (sum(out_juv_lfmm_pca$eig))) * 100, 2)

pca_dat_out_juv_lfmm <- as.data.frame(out_juv_lfmm_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_juv_lfmm.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Con1", "Con2", "Con3", "Hi1", "Hi2", "Hi3")))


ggplot(data = pca_dat_out_juv_lfmm) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_juv_lfmm[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_juv_lfmm[2], "%")) +
    coord_fixed(ratio = 1)
```

```{r}
rad_out_juv_lfmm_treat.filt <- vcfR2genind(my_vcf_out_juv_lfmm, strata = strata_juv, pop = c(rep("Control", 60),rep("CADO", 60)))

rad_out_juv_lfmm_treat.filt
```

```{r}
# Create color scheme
cols_juv_lfmm_treat = c("blue1","red")
```

```{r}
# Plot a PCA for the all outlier loci
x_out_juv_lfmm_treat <- scaleGen(rad_out_juv_lfmm_treat.filt, NA.method = "mean")
out_juv_lfmm_treat_pca <- dudi.pca(x_out_juv_lfmm_treat, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_juv_lfmm_treat <- round((out_juv_lfmm_treat_pca$eig / (sum(out_juv_lfmm_treat_pca$eig))) * 100, 2)

pca_dat_out_juv_lfmm_treat <- as.data.frame(out_juv_lfmm_treat_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_juv_lfmm_treat.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Control", "CADO")))


ggplot(data = pca_dat_out_juv_lfmm_treat) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv_lfmm_treat) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_juv_lfmm_treat[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_juv_lfmm_treat[2], "%")) +
    coord_fixed(ratio = 1)
```

```{r}
library(vegan)
```

```{r}
pc_scores <- out_juv_lfmm_treat_pca$li
```

```{r}
pop <- c(rep("Control", 60),rep("CADO", 60))
```

```{r}
permanova_result <- adonis2(pc_scores ~ pop, method = "euclidean")
print(permanova_result)
```

```{r}
pca_data <- data.frame(PC1 = pc_scores[,1], PC2 = pc_scores[,2], Treatment = pop)

ggplot(pca_data, aes(x = PC1, y = PC2, color = Treatment)) +
  geom_point() +
  stat_ellipse() +
  theme_minimal()
```

323 LFMM outliers

**Make genind object**

```{r}
# VCF file with juv_lfmm_0.6 SNP dataset
my_vcf_out_juv_lfmm_0.6 <- read.vcfR("outlierloci_juv_lfmm_0.6.recode.vcf")
```

```{r}
rad_out_juv_lfmm_0.6.filt <- vcfR2genind(my_vcf_out_juv_lfmm_0.6, strata = strata_juv, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))

rad_out_juv_lfmm_0.6.filt
```

**Make hierfstat object**

```{r}
hf_out_juv_lfmm_0.6.filt <- genind2hierfstat(rad_out_juv_lfmm_0.6.filt, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))
```

## Pairwise Fst  

```{r}
fst_out_juv_lfmm_0.6.mat <- pairwise.WCfst(hf_out_juv_lfmm_0.6.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_juv_lfmm_0.6_fst <- boot.ppfst(hf_out_juv_lfmm_0.6.filt,nboot = 1000)
boot5_out_juv_lfmm_0.6_fst <- boot.ppfst(hf_out_juv_lfmm_0.6.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_juv_lfmm_0.6_ll <- boot5_out_juv_lfmm_0.6_fst$ll
```

```{r}
gindF_out_juv_lfmm_0.6.fst.mat.triN <- as.matrix(fst_out_juv_lfmm_0.6.mat)
colnames(gindF_out_juv_lfmm_0.6.fst.mat.triN) <- pop_order_juv
rownames(gindF_out_juv_lfmm_0.6.fst.mat.triN) <- pop_order_juv
```

```{r}
meltedN_out_juv_lfmm_0.6 <- reshape2::melt(gindF_out_juv_lfmm_0.6.fst.mat.triN, na.rm =TRUE)
round(gindF_out_juv_lfmm_0.6.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_juv_lfmm_0.6$value)
```

```{r}
new_names <- c(
  "Con1" = "Control 1", "Con2" = "Control 2", "Con3" = "Control 3",
  "Hi1" = "Stress 1", "Hi2" = "Stress 2", "Hi3" = "Stress 3"
)

# Apply the new names to your melted data
meltedN_out_juv_lfmm_0.6$Var1 <- factor(new_names[as.character(meltedN_out_juv_lfmm_0.6$Var1)], levels = new_names)
meltedN_out_juv_lfmm_0.6$Var2 <- factor(new_names[as.character(meltedN_out_juv_lfmm_0.6$Var2)], levels = new_names)

#Plotting Pairwise fst
out_juv_lfmm_0.6 <- ggplot(data = meltedN_out_juv_lfmm_0.6, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "#25B396", name="FST")  +
  labs( x = "Treatment", y = "Treatment") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 10, hjust = 1),axis.text.y = element_text(size = 10)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =10), legend.title = element_text(size =10)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_juv_lfmm_0.6
```

## PCA 

```{r}
library("PCAviz")  #Visualizing output of PCA
library("cowplot") #Used with PCAviz
```

```{r}
# Create color scheme
cols_juv_lfmm_0.6 = c("blue1","lightblue","darkblue","red","red3","darkred")
```

```{r}
# Plot a PCA for the all outlier loci
x_out_juv_lfmm_0.6 <- scaleGen(rad_out_juv_lfmm_0.6.filt, NA.method = "mean")
out_juv_lfmm_0.6_pca <- dudi.pca(x_out_juv_lfmm_0.6, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_juv_lfmm_0.6 <- round((out_juv_lfmm_0.6_pca$eig / (sum(out_juv_lfmm_0.6_pca$eig))) * 100, 2)

pca_dat_out_juv_lfmm_0.6 <- as.data.frame(out_juv_lfmm_0.6_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_juv_lfmm_0.6.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Con1", "Con2", "Con3", "Hi1", "Hi2", "Hi3")))


ggplot(data = pca_dat_out_juv_lfmm_0.6) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_juv_lfmm_0.6[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_juv_lfmm_0.6[2], "%")) +
    coord_fixed(ratio = 1)
```

```{r}
rad_out_juv_lfmm_0.6_treat.filt <- vcfR2genind(my_vcf_out_juv_lfmm_0.6, strata = strata_juv, pop = c(rep("Control", 60),rep("CADO", 60)))

rad_out_juv_lfmm_0.6_treat.filt
```

```{r}
# Create color scheme
cols_juv_lfmm_0.6_treat = c("#56B4E9","orange")
```

```{r}
# Plot a PCA for the all outlier loci
x_out_juv_lfmm_0.6_treat <- scaleGen(rad_out_juv_lfmm_0.6_treat.filt, NA.method = "mean")
out_juv_lfmm_0.6_treat_pca <- dudi.pca(x_out_juv_lfmm_0.6_treat, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_juv_lfmm_0.6_treat <- round((out_juv_lfmm_0.6_treat_pca$eig / (sum(out_juv_lfmm_0.6_treat_pca$eig))) * 100, 2)

pca_dat_out_juv_lfmm_0.6_treat <- as.data.frame(out_juv_lfmm_0.6_treat_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_juv_lfmm_0.6_treat.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Control", "CADO")))


ggplot(data = pca_dat_out_juv_lfmm_0.6_treat) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv_lfmm_0.6_treat) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_juv_lfmm_0.6_treat[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_juv_lfmm_0.6_treat[2], "%")) +
    coord_fixed(ratio = 1)
```

```{r}
pc_scores <- out_juv_lfmm_0.6_treat_pca$li
```

```{r}
pop <- c(rep("Control", 60),rep("CADO", 60))
```

```{r}
permanova_result <- adonis2(pc_scores ~ pop, method = "euclidean")
print(permanova_result)
```

```{r}
pca_data <- data.frame(PC1 = pc_scores[,1], PC2 = pc_scores[,2], Treatment = pop)


# Define new treatment names
new_treatment_names <- c("Control" = "Control",
                         "CADO" = "Stress")

# Define custom colors
custom_colors <- c("Control" = "#56B4E9", "CADO" = "orange")

# Extract p-value from PERMANOVA result
p_value <- permanova_result$`Pr(>F)`[1]

eigenvalues <- eigen(cov(pca_data[, c("PC1", "PC2")]))$values

# Calculate proportion of variance explained
pve <- eigenvalues / sum(eigenvalues) * 100

# Create the plot with proportion of variance on axes
ggplot(pca_data, aes(x = PC1, y = PC2, color = Treatment)) +
  geom_point() +
  stat_ellipse() +
  theme_minimal() +
  scale_color_manual(values = custom_colors, labels = new_treatment_names) +
  xlab(paste0("PC1 (", round(pve[1], 2), "%)")) +
  ylab(paste0("PC2 (", round(pve[2], 2), "%)")) +
  annotate("text", x = -Inf, y = -Inf, 
           label = paste("PERMANOVA\n p =", format(p_value, digits = 2)),
           hjust = -0.1, vjust = -1)
```

```{r}
pca_loadings_out_juv_pc1 <- out_juv_lfmm_0.6_treat_pca$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  dplyr::arrange(desc(Comp1))

pca_loadings_out_juv_pc2 <- out_juv_lfmm_0.6_treat_pca$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  dplyr::arrange(desc(Comp2))
```


## Correspondence analysis-based RDA

```{r}
# Run a CA on the genind object
obj_out_juv <- genind2genpop(rad_out_juv_lfmm_0.6.filt)
ca1_out_juv <- dudi.coa(tab(obj_out_juv), scannf=FALSE, nf=10)
eig_ca1_out_juv <- round((ca1_out_juv$eig / (sum(ca1_out_juv$eig))) * 100, 2)
out_ca_out_juv <- ca1_out_juv$li[,1:2]
rownames(out_ca_out_juv) <- popNames(rad_out_juv_lfmm_0.6.filt)
```

```{r}
out_ca_out_juv <- out_ca_out_juv %>%
  rownames_to_column(var = "Treatment") %>%
  mutate(Treatment = factor(Treatment, levels = c("Con1", "Con2", "Con3", "Hi1", "Hi2", "Hi3"))) %>%
  dplyr::select(Treatment, Axis1, Axis2)
```

```{r}
# Plot the CA
ggplot(out_ca_out_juv, aes(x = Axis1, y = Axis2, colour = Treatment)) +
  geom_point(size = 6, alpha = 0.8) +
  scale_colour_manual(values = cols_juv_lfmm_0.6,
                      name = "") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14)) +
  labs(x = paste("", "Axis 1: ", eig_ca1_out_juv[1], "%"), y = paste("", "Axis 2: ", eig_ca1_out_juv[2], "%"))
```


```{r}
ca_loadings_out_juv_ca1 <- ca1_out_juv$co %>%
  as_tibble(rownames = "allele") %>%
  extract(allele, c("locus","chrom", "pos", "allele"), "(NC_(\\d+)_1_(\\d+)).(\\d)") %>%
  mutate(pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, allele, Comp1, Comp2) %>%
  mutate(Comp1 = Comp1^2, Comp2 = Comp2^2) %>%
  dplyr::arrange(desc(Comp1))
```

Making a file with loci chromosome and position info saved with the geneID

```{bash}
bedtools intersect -a <(awk '{print $1 "\t" $2-1 "\t" $2 "\t" $1"_"$2}' outliers_juv_lfmm2_0.6.loci.txt) -b sorted.ref3.0.gene.bed -wb | 
awk -F'\t' '{split($NF,a,"gene="); split(a[2],b,";"); print $1 "\t" $3 "\t" b[1]}' | 
sort -u > outliers_juv_lfmm2_0.6.with_genes.txt
```


## Making bed files to look for overlap between candidate larval and juvenile loci 

First extract chromosome and 

```{bash}
awk '{print $1 "\t" $2-1 "\t" $2}' outliers_juv_lfmm2_0.6.loci.txt | bedtools sort -i - > outliers_juv_lfmm2_06.bed

bedtools slop -b 10 -i outliers_juv_lfmm2_06.bed -g reference.fasta.fai | bedtools merge -i - > outliers_juv_lfmm2_06.intervals.bed
```

## Identifying genes for juvenile candidate loci

```{bash}
bedtools intersect -wb -a outliers_juv_lfmm2_06.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > Sig.gene.juv.CADO.LOC
```

194 genes 


## Gene Ontology with TopGO

I have a gaf from NCBI with the GO terms for genes in the eastern oyster genome. I'm going to take this file and extract the Gene IDs and associated GO terms.

In terminal:

```{bash}
wget https://ftp.ncbi.nlm.nih.gov/genomes/all/GCF/002/022/765/GCF_002022765.2_C_virginica-3.0/GCF_002022765.2_C_virginica-3.0_gene_ontology.gaf.gz

gunzip GCF_002022765.2_C_virginica-3.0_gene_ontology.gaf.gz

mv GCF_002022765.2_C_virginica-3.0_gene_ontology.gaf gene_ontology.gaf
```

Making a new .tab file with just the gene ID and GO IDs (3rd and 5th columns) renamed "gene_name" and "GeneOntologyIDs". The gene_name column will have one unique observation per row and all associated GO terms will be saved in the same row, with values separaetd by ;

```{bash}
awk 'NR>9 {print $3 "\t" $5}' gene_ontology.gaf | 
sed '1i gene_name\tGeneOntologyIDs' | 
sort | awk '
NR==1 {print; next}
$1!=prev {
if (NR>2) print prev "\t" goids;
prev=$1;
goids=$2;
next
}
{
goids=goids "; " $2
}
END {print prev "\t" goids}
' > gene2go_full.tab
```

Gene IDs for the candidate loci are saved to `Sig.gene.juv.CADO.LOC`

I'm making another file with the gene IDs and GO IDs for the candidate loci:

```{bash}
grep -F -f Sig.gene.juv.CADO.LOC gene2go_full.tab > juv_cand_genes_GO.tab

# Add header
awk 'BEGIN {print "gene_name\tGeneOntologyIDs"} {print}' juv_cand_genes_GO.tab > juv_cand_genes_GO_head.tab
```

In R:

```{r}
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("topGO")
```

```{r setup, include=FALSE}
library(topGO)
```

Read in gene2go information 

Old file names: all_genes (gene2go_full.tab); candidate genes (gene2go.tab)

```{r}
all_gene2go<-read.delim("gene2go_full.tab", sep="\t")
out_gene2go<-read.delim("juv_cand_genes_GO_head.tab", sep="\t")
```

Make list of genes for input to TopGO

```{r}
cand_genes <- as.character(unique(out_gene2go$gene_name))
all_genes <- as.character(all_gene2go$gene_name)
out_GeneList <- factor(as.integer(all_genes %in% cand_genes))
names(out_GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test  

Read in gene-to-go-mappings
```{r}
gene2go_topgo<-readMappings("gene2go_full.tab", IDsep=";", sep="\t")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_BP_pi <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_BP_FE_pi <- runTest(GO_OUT_BP_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_BP_En_pi <- GenTable(GO_OUT_BP_pi, Fisher = GO_OUT_BP_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_BP_En_pi$Fisher<-as.numeric(GO_OUT_BP_En_pi$Fisher)
GO_OUT_BP_En_sig_pi<-GO_OUT_BP_En_pi[GO_OUT_BP_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_BP_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_BP_En_sig_pi$GO.ID <- trimws(GO_OUT_BP_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_BP_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_BP_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_BP_En_sig_gene_pi$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_CC_pi <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_CC_FE_pi <- runTest(GO_OUT_CC_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_CC_En_pi <- GenTable(GO_OUT_CC_pi, Fisher = GO_OUT_CC_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_CC_En_pi$Fisher<-as.numeric(GO_OUT_CC_En_pi$Fisher)
GO_OUT_CC_En_sig_pi<-GO_OUT_CC_En_pi[GO_OUT_CC_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_CC_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_CC_En_sig_pi$GO.ID <- trimws(GO_OUT_CC_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_CC_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_CC_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_CC_En_sig_gene_pi$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_MF_pi <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_MF_FE_pi <- runTest(GO_OUT_MF_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_MF_En_pi <- GenTable(GO_OUT_MF_pi, Fisher = GO_OUT_MF_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_MF_En_pi$Fisher<-as.numeric(GO_OUT_MF_En_pi$Fisher)
GO_OUT_MF_En_sig_pi<-GO_OUT_MF_En_pi[GO_OUT_MF_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_MF_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_MF_En_sig_pi$GO.ID <- trimws(GO_OUT_MF_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_MF_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_MF_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_MF_En_sig_gene_pi$ontology <- "Molecular Functions"
```

### Join ontologies

Bind `GO_OUT_BP_En_sig_gene_pi`, `GO_OUT_CC_En_sig_gene_pi`, and `GO_OUT_MF_En_sig_gene_pi` to make a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_OUT_En_sig_gene_pi <- rbind(GO_OUT_BP_En_sig_gene_pi, GO_OUT_CC_En_sig_gene_pi, GO_OUT_MF_En_sig_gene_pi)
```

Merge `pi_bed_filt` with `GO_PEVE_En_sig_gene` and calculate proportion of significant genes
```{r}
merged_out_genes <- GO_OUT_En_sig_gene_pi %>%
  mutate(prop.sig.genes = Significant/Annotated)

# Write to csv 
write.csv(merged_out_genes, "Juv_OUT_GO_en_sig_gene.csv")
```

Plot - need to work out 
```{r}
GO_plot_pi<-ggplot(merged_out_genes, aes(x = Term, y = -log(Fisher), size = prop.sig.genes, fill = -log(Fisher))) +
  expand_limits(y = 1.5) +
  ylim(1, 7.25) +
  #Add horizontal lines with a single aesthetic value
  geom_hline(yintercept = -log10(0.01), linetype = "longdash", colour = "black", linewidth = .6) +
  geom_hline(yintercept = -log10(0.001), linetype = "solid", colour = "black", linewidth = .6) +
  geom_point(shape = 21) + 
  scale_size(range = c(2, 12)) + 
  scale_fill_continuous(low = "#1AD3D1FF", high = "#4686FBFF") +
  xlab('') + 
  ylab('Enrichment score') +
  #labs(caption = 'Cut-off lines at p=0.01 and p=0.001') +
  theme_bw() +
  facet_grid(vars(ontology), scales = "free", space = "free_y") +
  coord_flip(); GO_plot_pi
```







145 LFMM outliers

**Make genind object**

```{r}
# VCF file with juv_lfmm_0.64 SNP dataset
my_vcf_out_juv_lfmm_0.64 <- read.vcfR("outlierloci_juv_lfmm_0.64.recode.vcf")
```

```{r}
rad_out_juv_lfmm_0.64.filt <- vcfR2genind(my_vcf_out_juv_lfmm_0.64, strata = strata_juv, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))

rad_out_juv_lfmm_0.64.filt
```

**Make hierfstat object**

```{r}
hf_out_juv_lfmm_0.64.filt <- genind2hierfstat(rad_out_juv_lfmm_0.64.filt, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))
```

## Pairwise Fst  

```{r}
fst_out_juv_lfmm_0.64.mat <- pairwise.WCfst(hf_out_juv_lfmm_0.64.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_juv_lfmm_0.64_fst <- boot.ppfst(hf_out_juv_lfmm_0.64.filt,nboot = 1000)
boot5_out_juv_lfmm_0.64_fst <- boot.ppfst(hf_out_juv_lfmm_0.64.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_juv_lfmm_0.64_ll <- boot5_out_juv_lfmm_0.64_fst$ll
```

```{r}
gindF_out_juv_lfmm_0.64.fst.mat.triN <- as.matrix(fst_out_juv_lfmm_0.64.mat)
colnames(gindF_out_juv_lfmm_0.64.fst.mat.triN) <- pop_order_juv
rownames(gindF_out_juv_lfmm_0.64.fst.mat.triN) <- pop_order_juv
```

```{r}
meltedN_out_juv_lfmm_0.64 <- melt(gindF_out_juv_lfmm_0.64.fst.mat.triN, na.rm =TRUE)
round(gindF_out_juv_lfmm_0.64.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_juv_lfmm_0.64$value)
```

```{r}
#Plotting Pairwise fst
out_juv_lfmm_0.64 <- ggplot(data = meltedN_out_juv_lfmm_0.64, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) juv_lfmm outlier SNPs", atop(italic("N = 120, L = 145"), ""))))+
  labs( x = "Treatment", y = "Treatment") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_juv_lfmm_0.64
```

## PCA 

```{r}
library("PCAviz")  #Visualizing output of PCA
library("cowplot") #Used with PCAviz
```

```{r}
# Create color scheme
cols_juv_lfmm_0.64 = c("blue1","lightblue","darkblue","red","red3","darkred")
```

```{r}
# Plot a PCA for the all outlier loci
x_out_juv_lfmm_0.64 <- scaleGen(rad_out_juv_lfmm_0.64.filt, NA.method = "mean")
out_juv_lfmm_0.64_pca <- dudi.pca(x_out_juv_lfmm_0.64, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_juv_lfmm_0.64 <- round((out_juv_lfmm_0.64_pca$eig / (sum(out_juv_lfmm_0.64_pca$eig))) * 100, 2)

pca_dat_out_juv_lfmm_0.64 <- as.data.frame(out_juv_lfmm_0.64_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_juv_lfmm_0.64.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Con1", "Con2", "Con3", "Hi1", "Hi2", "Hi3")))


ggplot(data = pca_dat_out_juv_lfmm_0.64) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_juv_lfmm_0.64[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_juv_lfmm_0.64[2], "%")) +
    coord_fixed(ratio = 1)
```

30 LFMM outliers

**Make genind object**

```{r}
# VCF file with juv_lfmm_0.7 SNP dataset
my_vcf_out_juv_lfmm_0.7 <- read.vcfR("outlierloci_juv_lfmm_0.7.recode.vcf")
```

```{r}
rad_out_juv_lfmm_0.7.filt <- vcfR2genind(my_vcf_out_juv_lfmm_0.7, strata = strata_juv, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))

rad_out_juv_lfmm_0.7.filt
```

**Make hierfstat object**

```{r}
hf_out_juv_lfmm_0.7.filt <- genind2hierfstat(rad_out_juv_lfmm_0.7.filt, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))
```

## Pairwise Fst  

```{r}
fst_out_juv_lfmm_0.7.mat <- pairwise.WCfst(hf_out_juv_lfmm_0.7.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_juv_lfmm_0.7_fst <- boot.ppfst(hf_out_juv_lfmm_0.7.filt,nboot = 1000)
boot5_out_juv_lfmm_0.7_fst <- boot.ppfst(hf_out_juv_lfmm_0.7.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_juv_lfmm_0.7_ll <- boot5_out_juv_lfmm_0.7_fst$ll
```

```{r}
gindF_out_juv_lfmm_0.7.fst.mat.triN <- as.matrix(fst_out_juv_lfmm_0.7.mat)
colnames(gindF_out_juv_lfmm_0.7.fst.mat.triN) <- pop_order_juv
rownames(gindF_out_juv_lfmm_0.7.fst.mat.triN) <- pop_order_juv
```

```{r}
meltedN_out_juv_lfmm_0.7 <- melt(gindF_out_juv_lfmm_0.7.fst.mat.triN, na.rm =TRUE)
round(gindF_out_juv_lfmm_0.7.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_juv_lfmm_0.7$value)
```

```{r}
#Plotting Pairwise fst
out_juv_lfmm_0.7 <- ggplot(data = meltedN_out_juv_lfmm_0.7, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) juv_lfmm_0.7 outlier SNPs", atop(italic("N = 120, L = 30"), ""))))+
  labs( x = "Treatment", y = "Treatment") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_juv_lfmm_0.7
```

## PCA 

```{r}
library("PCAviz")  #Visualizing output of PCA
library("cowplot") #Used with PCAviz
```

```{r}
# Create color scheme
cols_juv_lfmm_0.7 = c("blue1","lightblue","darkblue","red","red3","darkred")
```

```{r}
# Plot a PCA for the all outlier loci
x_out_juv_lfmm_0.7 <- scaleGen(rad_out_juv_lfmm_0.7.filt, NA.method = "mean")
out_juv_lfmm_0.7_pca <- dudi.pca(x_out_juv_lfmm_0.7, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_juv_lfmm_0.7 <- round((out_juv_lfmm_0.7_pca$eig / (sum(out_juv_lfmm_0.7_pca$eig))) * 100, 2)

pca_dat_out_juv_lfmm_0.7 <- as.data.frame(out_juv_lfmm_0.7_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_juv_lfmm_0.7.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Con1", "Con2", "Con3", "Hi1", "Hi2", "Hi3")))


ggplot(data = pca_dat_out_juv_lfmm_0.7) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_juv_lfmm_0.7[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_juv_lfmm_0.7[2], "%")) +
    coord_fixed(ratio = 1)
```


pcaadapt + outflank outliers

**Make genind object**

```{r}
# VCF file with juv_db SNP dataset
my_vcf_out_juv_db <- read.vcfR("outlierloci_juv_db.recode.vcf")
```

```{r}
rad_out_juv_db.filt <- vcfR2genind(my_vcf_out_juv_db, strata = strata_juv, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))

rad_out_juv_db.filt
```

**Make hierfstat object**

```{r}
hf_out_juv_db.filt <- genind2hierfstat(rad_out_juv_db.filt, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))
```

## Pairwise Fst  

```{r}
fst_out_juv_db.mat <- pairwise.WCfst(hf_out_juv_db.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_out_juv_db_fst <- boot.ppfst(hf_out_juv_db.filt,nboot = 1000)
boot5_out_juv_db_fst <- boot.ppfst(hf_out_juv_db.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_out_juv_db_ll <- boot5_out_juv_db_fst$ll
```

```{r}
gindF_out_juv_db.fst.mat.triN <- as.matrix(fst_out_juv_db.mat)
colnames(gindF_out_juv_db.fst.mat.triN) <- pop_order_juv
rownames(gindF_out_juv_db.fst.mat.triN) <- pop_order_juv
```

```{r}
meltedN_out_juv_db <- melt(gindF_out_juv_db.fst.mat.triN, na.rm =TRUE)
round(gindF_out_juv_db.fst.mat.triN,4)
```

```{r}
summary(meltedN_out_juv_db$value)
```

```{r}
#Plotting Pairwise fst
out_juv_db <- ggplot(data = meltedN_out_juv_db, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) juv_db outlier SNPs", atop(italic("N = 120, L = 322"), ""))))+
  labs( x = "Treatment", y = "Treatment") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
out_juv_db
```

## PCA 

```{r}
library("PCAviz")  #Visualizing output of PCA
library("cowplot") #Used with PCAviz
```

```{r}
# Create color scheme
cols_juv_db = c("blue1","lightblue","darkblue","red","red3","darkred")
```

```{r}
# Plot a PCA for the all outlier loci
x_out_juv_db <- scaleGen(rad_out_juv_db.filt, NA.method = "mean")
out_juv_db_pca <- dudi.pca(x_out_juv_db, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_out_juv_db <- round((out_juv_db_pca$eig / (sum(out_juv_db_pca$eig))) * 100, 2)

pca_dat_out_juv_db <- as.data.frame(out_juv_db_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_out_juv_db.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Con1", "Con2", "Con3", "Hi1", "Hi2", "Hi3")))


ggplot(data = pca_dat_out_juv_db) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_out_juv_db[1], "%"), y = paste("", "Axis 2: ", eig_percent_out_juv_db[2], "%")) +
    coord_fixed(ratio = 1)
```


### All SNPs

```{r}
# VCF file with juv SNP dataset
my_vcf_all_juv <- read.vcfR("../SNP.Juvdp10mafMIp9g9nDNAmaf052A.recode.vcf")
```

```{r}
rad_all_juv.filt <- vcfR2genind(my_vcf_all_juv, strata = strata_juv, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))

rad_all_juv.filt
```

**Make hierfstat object**

```{r}
hf_all_juv.filt <- genind2hierfstat(rad_all_juv.filt, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))
```

## Pairwise Fst  

```{r}
fst_all_juv.mat <- pairwise.WCfst(hf_all_juv.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_all_juv_fst <- boot.ppfst(hf_all_juv.filt,nboot = 1000)
boot5_all_juv_fst <- boot.ppfst(hf_all_juv.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_all_juv_ll <- boot5_all_juv_fst$ll
```

```{r}
gindF_all_juv.fst.mat.triN <- as.matrix(fst_all_juv.mat)
colnames(gindF_all_juv.fst.mat.triN) <- pop_order_juv
rownames(gindF_all_juv.fst.mat.triN) <- pop_order_juv
```

```{r}
meltedN_all_juv <- melt(gindF_all_juv.fst.mat.triN, na.rm =TRUE)
round(gindF_all_juv.fst.mat.triN,4)
```

```{r}
summary(meltedN_all_juv$value)
```

```{r}
#Plotting Pairwise fst
all_juv <- ggplot(data = meltedN_all_juv, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) juv alllier SNPs", atop(italic("N = 120, L = 346"), ""))))+
  labs( x = "Treatment", y = "Treatment") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
all_juv
```

## PCA 

```{r}
library("PCAviz")  #Visualizing allput of PCA
library("cowplot") #Used with PCAviz
```

```{r}
# Create color scheme
cols_juv = c("blue1","lightblue","darkblue","red","red3","darkred")
```

```{r}
# Plot a PCA for all loci
x_all_juv <- scaleGen(rad_all_juv.filt, NA.method = "mean")
all_juv_pca <- dudi.pca(x_all_juv, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_all_juv <- round((all_juv_pca$eig / (sum(all_juv_pca$eig))) * 100, 2)

pca_dat_all_juv <- as.data.frame(all_juv_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_all_juv.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Con1", "Con2", "Con3", "Hi1", "Hi2", "Hi3")))


ggplot(data = pca_dat_all_juv) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv) +
    theme_minimal() +
    labs(x = paste("", "Axis 1: ", eig_percent_all_juv[1], "%"), y = paste("", "Axis 2: ", eig_percent_all_juv[2], "%")) +
    coord_fixed(ratio = 1)
```

### Neutral SNPs

$ vcftools --vcf ../SNP.Juvdp10mafMIp9g9nDNAmaf052A.recode.vcf --exclude-positions outliers_juv_lfmm2_0.6.loci.txt --recode --recode-INFO-all --out neutralloci_juv 

```{r}
# VCF file with juv SNP dataset
my_vcf_neut_juv <- read.vcfR("neutralloci_juv.recode.vcf")
```

```{r}
rad_neut_juv.filt <- vcfR2genind(my_vcf_neut_juv, strata = strata_juv, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))

rad_neut_juv.filt
```

**Make hierfstat object**

```{r}
hf_neut_juv.filt <- genind2hierfstat(rad_neut_juv.filt, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))
```

## Pairwise Fst  

```{r}
fst_neut_juv.mat <- pairwise.WCfst(hf_neut_juv.filt)
```

```{r}
# get bootstrap confidence values for Fst
boot_neut_juv_fst <- boot.ppfst(hf_neut_juv.filt,nboot = 1000)
boot5_neut_juv_fst <- boot.ppfst(hf_neut_juv.filt,nboot = 1000,quant = 0.5)
```

```{r}
Fst_neut_juv_ll <- boot5_neut_juv_fst$ll
```

```{r}
gindF_neut_juv.fst.mat.triN <- as.matrix(fst_neut_juv.mat)
colnames(gindF_neut_juv.fst.mat.triN) <- pop_order_juv
rownames(gindF_neut_juv.fst.mat.triN) <- pop_order_juv
```

```{r}
meltedN_neut_juv <- melt(gindF_neut_juv.fst.mat.triN, na.rm =TRUE)
round(gindF_neut_juv.fst.mat.triN,4)
```

```{r}
summary(meltedN_neut_juv$value)
```

```{r}
#Plotting Pairwise fst
neut_juv <- ggplot(data = meltedN_neut_juv, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "blue1", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) juv Neutral SNPs", atop(italic("N = 120, L = 47,313"), ""))))+
  labs( x = "Treatment", y = "Treatment") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 12)) + 
  theme(axis.title = element_text(size = 12),legend.text = element_text(size =14), legend.title = element_text(size =14)) +
  theme(plot.title = element_text(size = 14)) +
coord_fixed()
neut_juv
```



## PCA 

```{r}
rad_neut_juv_treat.filt <- vcfR2genind(my_vcf_neut_juv, strata = strata_juv, pop = c(rep("Control", 60),rep("Stress", 60)))

rad_neut_juv_treat.filt
```

```{r}
library("PCAviz")  #Visualizing neutput of PCA
library("cowplot") #Used with PCAviz
```

```{r}
# Plot a PCA for neut loci
x_neut_juv <- scaleGen(rad_neut_juv_treat.filt, NA.method = "mean")
neut_juv_pca <- dudi.pca(x_neut_juv, center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig_percent_neut_juv <- round((neut_juv_pca$eig / (sum(neut_juv_pca$eig))) * 100, 2)

pca_dat_neut_juv <- as.data.frame(neut_juv_pca$li) %>%
                    rownames_to_column(var = "ind") %>%
                    cbind(pop = pop(rad_neut_juv_treat.filt)) %>%
                    mutate(pop = factor(pop, levels = c("Control", "Stress")))

# Create color scheme
cols_juv <- c("#56B4E9", "orange")

# Plot a PCA for neut loci
ggplot(data = pca_dat_neut_juv) +
    geom_point(aes(x = Axis1, y = Axis2, colour = pop), size = 3, alpha = 0.8) +
    scale_colour_manual(values = cols_juv) +
    theme_bw() + 
    labs(x = paste("", "Axis 1: ", eig_percent_neut_juv[1], "%"), y = paste("", "Axis 2: ", eig_percent_neut_juv[2], "%")) +
    coord_fixed(ratio = 1) +
  theme(legend.title = element_blank())
ggsave("juv_pca_neutral.pdf", width = 7, height = 6, units = "in")
```


## Manhattan Plot

```{r}
# Calculate basic stats for each population
pop_stats_out_juv <- rad_out_juv_lfmm_0.6.filt %>%
  genind2hierfstat() %>%
  basic.stats()

global_stats_out_juv <- pop_stats_out_juv$perloc %>%
  rownames_to_column(var = "locus") %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  as_tibble() %>%
  mutate(chrom = as.integer(chrom), pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, Ho, Hs, Fst, Fis)

# Adding columns to specify loci as outliers 
global_stats_out_juv$juv_sel_outlier <- TRUE
global_stats_out_juv$shared_sel_outlier <- FALSE
```

```{r}
# Calculate basic stats for each population
pop_stats_neut_juv <- rad_neut_juv.filt %>%
  genind2hierfstat() %>%
  basic.stats()

global_stats_neut_juv <- pop_stats_neut_juv$perloc %>%
  rownames_to_column(var = "locus") %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  as_tibble() %>%
  mutate(chrom = as.integer(chrom), pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, Ho, Hs, Fst, Fis)

# Adding columns to specify loci as outliers 
global_stats_neut_juv$juv_sel_outlier <- FALSE
global_stats_out_juv$shared_sel_outlier <- FALSE
```

Combining the tables together 

```{r}
combined_df_juv <- bind_rows(global_stats_out_juv, global_stats_neut_juv)
```

```{r}
# Change naming of chromosomes
chr_labs <- c("Chr 1", "Chr 2", "Chr 3", "Chr 4", "Chr 5", "Chr 6", "Chr 7", "Chr 8", "Chr 9","Chr 10")
names(chr_labs) <- c(35780, 35781,35782, 35783, 35784, 35785, 35786, 35787, 35788, 35789)

# Create the plot
p <- combined_df_juv %>%
  mutate(pos_mb = pos / 1000000) %>%
  ggplot(aes(x = pos_mb, y = Fst)) +
  # Add grey points for non-outliers
  geom_point(color = "grey", size = 1) +
  # Add orange points for outliers
  geom_point(data = . %>% filter(juv_sel_outlier == TRUE), 
             aes(color = "Post-settlement candidate loci"), size = 2) +
  facet_wrap(~chrom, scales = "free_x", nrow = 1, labeller = labeller(chrom = chr_labs)) +
  theme_minimal() +
  scale_color_manual(values = c("Post-settlement candidate loci" = "cyan4"),
                     name = NULL) +
  labs(x = "Position (Mb)", y = expression(F[ST])) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),
        strip.text.x = element_text(size = 10, face = "bold"),
        panel.spacing = unit(0.1, "lines"),
        legend.position = "right")
ggsave("juv_global_fst.pdf", width = 12.5, height = 5, units = "in")

# Display the plot
print(p)
```

Repeating with shared loci between larvae and juveniles 

```{r}
# VCF file with juv SNP dataset
my_vcf_shared <- read.vcfR("~/RISG_CADO/Juv_Larv_Comparison/CADO_shared_0.6_FDR05_intervals.recode.vcf")
```

```{r}
rad_out_shared.filt <- vcfR2genind(my_vcf_shared, strata = strata_juv, pop = c(rep("Con1", 20),rep("Con2", 20),rep("Con3", 20),rep("Hi1", 20),rep("Hi2", 20), rep("Hi3",20)))

rad_out_shared.filt
```

```{r}
# Calculate basic stats for each population
pop_stats_shared <- rad_out_shared.filt %>%
  genind2hierfstat() %>%
  basic.stats()

global_stats_shared <- pop_stats_shared$perloc %>%
  rownames_to_column(var = "locus") %>%
  extract(locus, c("chrom", "pos"), "NC_(\\d+)_1_(\\d+)", remove = FALSE) %>%
  as_tibble() %>%
  mutate(chrom = as.integer(chrom), pos = as.integer(pos)) %>%
  dplyr::select(locus, chrom, pos, Ho, Hs, Fst, Fis)

# Adding columns to specify loci as outliers 
global_stats_shared$juv_sel_outlier <- FALSE
global_stats_shared$shared_sel_outlier <- TRUE
```

Combining the tables together 

```{r}
combined_df_shared <- bind_rows(global_stats_out_juv, global_stats_neut_juv, global_stats_shared)
```

```{r}
# Change naming of chromosomes
chr_labs <- c("Chr 1", "Chr 2", "Chr 3", "Chr 4", "Chr 5", "Chr 6", "Chr 7", "Chr 8", "Chr 9","Chr 10")
names(chr_labs) <- c(35780, 35781,35782, 35783, 35784, 35785, 35786, 35787, 35788, 35789)

# Create the plot
p <- combined_df_shared %>%
  mutate(pos_mb = pos / 1000000) %>%
  ggplot(aes(x = pos_mb, y = Fst)) +
  # Add grey points for non-outliers
  geom_point(color = "grey", size = 1) +
  # Add orange points for juv_sel_outlier
  geom_point(data = . %>% filter(juv_sel_outlier == TRUE), 
             aes(color = "Post-settlement candidate loci"), size = 2) +
  # Add green points for shared_sel_outlier
  geom_point(data = . %>% filter(shared_sel_outlier == TRUE),
             aes(color = "Shared candidate loci"), size = 2) +
  facet_wrap(~chrom, scales = "free_x", nrow = 1, labeller = labeller(chrom = chr_labs)) +
  theme_minimal() +
  scale_color_manual(values = c("Post-settlement candidate loci" = "orange",
                                "Shared candidate loci" = "green4"),
                     name = NULL) +
  labs(x = "Position (Mb)", y = expression(F[ST])) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),
        strip.text.x = element_text(size = 10, face = "bold"),
        panel.spacing = unit(0.1, "lines"),
        legend.position = "bottom")

# Display the plot
print(p)
```


## All SNPs

Should go back and look at these across the genome as well to see chromosomal locations of selection in the Stress treatment compared to the Control.

### Nucleotide diversity

In PATH: `/home/azyck/RISG_CADO/juv_ddocent/Juv_Selection_Analysis`

```{bash}
mkdir nucleotide_diversity
```

```{bash}
cd nucleotide diversity
```

I have a file from the dDocent run with all sample names and replicate designations. I'm going to use this to split the VCF into replicates for both the Control and Stress treatments.

```{bash}
cp ../../popmap .
```

```{bash}
# Control samples
awk -F"\t" '$2 ~ /^Con1/ {print $1}' popmap > Control1_samples.txt
head Control1_samples.txt 

awk -F"\t" '$2 ~ /^Con2/ {print $1}' popmap > Control2_samples.txt
head Control2_samples.txt 

awk -F"\t" '$2 ~ /^Con3/ {print $1}' popmap > Control3_samples.txt
head Control3_samples.txt 

# Stress samples
awk -F"\t" '$2 ~ /^Hi1/ {print $1}' popmap > Stress1_samples.txt
head Stress1_samples.txt

awk -F"\t" '$2 ~ /^Hi2/ {print $1}' popmap > Stress2_samples.txt
head Stress2_samples.txt

awk -F"\t" '$2 ~ /^Hi3/ {print $1}' popmap > Stress3_samples.txt
head Stress3_samples.txt
```

Now, filter the vcf so it only keeps the Control samples

```{bash}
vcftools --vcf ../SNP.Juvdp10mafMIp9g9nDNAmaf052A.recode.vcf --keep Control1_samples.txt --recode --recode-INFO-all --out Control1

vcftools --vcf ../SNP.Juvdp10mafMIp9g9nDNAmaf052A.recode.vcf --keep Control2_samples.txt --recode --recode-INFO-all --out Control2

vcftools --vcf ../SNP.Juvdp10mafMIp9g9nDNAmaf052A.recode.vcf --keep Control3_samples.txt --recode --recode-INFO-all --out Control3
```

and Stress samples

```{bash}
vcftools --vcf ../SNP.Juvdp10mafMIp9g9nDNAmaf052A.recode.vcf --keep Stress1_samples.txt --recode --recode-INFO-all --out Stress1

vcftools --vcf ../SNP.Juvdp10mafMIp9g9nDNAmaf052A.recode.vcf --keep Stress2_samples.txt --recode --recode-INFO-all --out Stress2

vcftools --vcf ../SNP.Juvdp10mafMIp9g9nDNAmaf052A.recode.vcf --keep Stress3_samples.txt --recode --recode-INFO-all --out Stress3
```

**Calculate Nucleotide Diversity**

Use sliding window analysis for each group (10kb windows recommended):

```{bash}
# Control group π
vcftools --vcf Control1.recode.vcf --window-pi 10000 --out Control1_10kb

vcftools --vcf Control2.recode.vcf --window-pi 10000 --out Control2_10kb

vcftools --vcf Control3.recode.vcf --window-pi 10000 --out Control3_10kb

# Stress group π
vcftools --vcf Stress1.recode.vcf --window-pi 10000 --out Stress1_10kb

vcftools --vcf Stress2.recode.vcf --window-pi 10000 --out Stress2_10kb

vcftools --vcf Stress3.recode.vcf --window-pi 10000 --out Stress3_10kb
```

```{bash}
# Combining the replicates for each treatment into one, with Pop (replicate) and Group (treatment) columns added
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_VARIANTS\tPI\tPOP\tGROUP")  <(mawk '{if (NR>1)print $0 "\tCON1\tCONTROL"}' Control1_10kb.windowed.pi)   <(mawk '{if (NR>1)print $0 "\tCON2\tCONTROL"}' Control2_10kb.windowed.pi) <(mawk '{if (NR>1)print $0 "\tCON3\tCONTROL"}' Control3_10kb.windowed.pi) > Control.combined.windowed.pi

cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_VARIANTS\tPI\tPOP\tGROUP")  <(mawk '{if (NR>1)print $0 "\tHI1\tSTRESS"}' Stress1_10kb.windowed.pi)   <(mawk '{if (NR>1)print $0 "\tHI2\tSTRESS"}' Stress2_10kb.windowed.pi) <(mawk '{if (NR>1)print $0 "\tHI3\tSTRESS"}' Stress3_10kb.windowed.pi) > Stress.combined.windowed.pi

# Combining all into one meta file
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_VARIANTS\tPI\tPOP\tGROUP") <(mawk '{if (NR>1)print $0 "\tCON1\tCONTROL"}' Control1_10kb.windowed.pi)   <(mawk '{if (NR>1)print $0 "\tCON2\tCONTROL"}' Control2_10kb.windowed.pi) <(mawk '{if (NR>1)print $0 "\tCON3\tCONTROL"}' Control3_10kb.windowed.pi) <(mawk '{if (NR>1)print $0 "\tHI1\tSTRESS"}' Stress1_10kb.windowed.pi)   <(mawk '{if (NR>1)print $0 "\tHI2\tSTRESS"}' Stress2_10kb.windowed.pi) <(mawk '{if (NR>1)print $0 "\tHI3\tSTRESS"}' Stress3_10kb.windowed.pi) > all.windowed.pi
```

This next command merges overlapping intervals in the treatment π file and, for each merged region, calculates the mean number of variants and mean π, and keeps the first group label, producing a summary file

```{bash}
bedtools merge -i <(mawk '!/CHR/' Control.combined.windowed.pi | sort -k1,1 -k2,2n ) -c 4,5,7 -d -1 -o mean,mean,first > mean.Control.windowed.pi

bedtools merge -i <(mawk '!/CHR/' Stress.combined.windowed.pi | sort -k1,1 -k2,2n ) -c 4,5,7 -d -1 -o mean,mean,first > mean.Stress.windowed.pi
```

```{bash}
# Adding headers
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_VARIANTS\tPI\tGROUP") mean.Control.windowed.pi > mean.Control.windowed.pi.txt

cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_VARIANTS\tPI\tGROUP") mean.Stress.windowed.pi > mean.Stress.windowed.pi.txt
```

```{bash}
# Combining into one meta file
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_VARIANTS\tPI\tGROUP") mean.Control.windowed.pi mean.Stress.windowed.pi > mean.treat.windowed.pi
```

**Compare Diversity Between Treatments**

```{r}
prefix.I <- "../nucleotide_diversity/"
```

```{r}
pi.all.dataframe<-read.table(paste(prefix.I,"all.windowed.pi", sep=""), sep="\t", header=T)
pi.all.dataframe$GROUP <- factor(pi.all.dataframe$GROUP, levels=c("CONTROL", "STRESS" ))
pi.all.dataframe$POP <- gsub("^HI", "STR", pi.all.dataframe$POP)
pi.all.dataframe$POP <- factor(
  pi.all.dataframe$POP,
  levels = c("CON1", "CON2", "CON3", "STR1", "STR2", "STR3")
)


filename <- paste(prefix.I,"mean.treat.windowed.pi", sep="")
pi.treat.mean.dataframe<-read.table(filename, sep="\t", header=T)
pi.treat.mean.dataframe$GROUP <- factor(pi.treat.mean.dataframe$GROUP, levels=c("CONTROL", "STRESS" ))
```

Plotting

```{r}
bd.pi <-ggplot(pi.all.dataframe, aes(x=PI,y = POP))+
  #geom_violin(aes(color=GROUP,fill= GROUP)) +
  #stat_summary(fun=mean, geom="point", shape=23, size=4)+
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=0.005)+
  geom_boxplot(aes(fill=GROUP), width=0.75,outlier.shape = 23, outlier.color = "black")+
  geom_vline(xintercept =mean(pi.all.dataframe$PI),linetype = "dashed") +
  stat_summary(fun=mean, geom="point", shape=23, size=2)+
  scale_fill_manual(values=c("#56B4E9", "orange"))+
  scale_color_manual(values=c("#56B4E9", "orange"))+
  scale_y_discrete(limits = rev(levels(pi.all.dataframe$POP)))+
  #ylab("Population/Line")+
  labs(x=expression(pi)) +
  ggtitle("Control vs. Stress Nucleotide Diversity")+
  theme_classic() +
  theme(axis.title.y = element_blank()) + 
  theme(legend.position = "none")
bd.pi
```

```{r}
bc.pi <-ggplot(pi.treat.mean.dataframe, aes(x=PI,y = GROUP))+
  geom_violin(aes(color=GROUP,fill= GROUP)) +
  geom_boxplot(aes(fill=GROUP), width=0.1,outlier.shape = 23, outlier.color = "black")+
  stat_summary(fun=mean, geom="point", shape=23, size=2)+
  scale_fill_manual(values=c("#56B4E9", "orange"))+
  scale_color_manual(values=c("#56B4E9", "orange"))+
  scale_y_discrete(limits = rev(levels(pi.all.dataframe$GROUP)))+
  #ylab("Origin")+
  ggtitle("Mean Control vs. Stress Nucleotide Diversity")+
  theme_classic() +
  labs(x=expression(pi)) +
  theme(legend.position = c(0.85, 0.5), axis.title.y = element_blank(),axis.text.y=element_blank())
bc.pi
```

Running t-tests

```{r}
library(infer)
library(plotrix)
```

```{r}
pi.all.dataframe %>% t_test(formula = PI ~ GROUP, alternative = "two.sided",order = c("CONTROL", "STRESS"))
```

```{r}
pi.all.dataframe %>% t_test(formula = PI ~ GROUP, alternative = "greater", order = c("CONTROL", "STRESS"))
```

**Calculating summary statistics**

```{r}
group_by(pi.all.dataframe, GROUP) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(PI, na.rm = TRUE),
    sd = sd(PI, na.rm = TRUE),
    se=std.error(PI, na.rm = TRUE) )
```

```{r}
group_by(pi.all.dataframe, POP) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(PI, na.rm = TRUE),
    sd = sd(PI, na.rm = TRUE),
    se=std.error(PI, na.rm = TRUE) )
```

**Plotting PI across the genome**

```{r}
filename <- paste(prefix.I,"mean.Control.windowed.pi.txt", sep="")
cpi <- read.table(filename, header = TRUE)
filename <- paste(prefix.I,"mean.Stress.windowed.pi.txt", sep="")
spi <- read.table(filename, header = TRUE)

cpi <- cpi %>% dplyr::rename(CONTROL_PI= PI)
spi <- spi %>% dplyr::rename(STRESS_PI= PI)
c.s.pi <- plyr::join(cpi,spi, by =c("CHROM", "BIN_START"),type="full")


c.s.pi$DIFF <- c.s.pi$CONTROL_PI - c.s.pi$STRESS_PI
c.s.pi$BIN_START <- c.s.pi$BIN_START -1
```

```{r}
SNP.pi<-c(1:(nrow(c.s.pi)))
mydf.pi<-data.frame(SNP.pi,c.s.pi)
mydf.pi$CHR <- mydf.pi$CHROM
mydf.pi %>% 
  mutate(CHR = str_replace(CHR, "NC_035780.1", "1")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035781.1", "2")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035782.1", "3")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035783.1", "4")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035784.1", "5")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035785.1", "6")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035786.1", "7")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035787.1", "8")) %>%
  mutate(CHR = str_replace(CHR, "NC_035788.1", "9")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035789.1", "10"))  -> mydf.pi
mydf.pi$CHR <- as.numeric(mydf.pi$CHR)  

m2.pi <-ggman::ggman(mydf.pi,chrom="CHR",bp="BIN_START",pvalue="CONTROL_PI",snp="SNP.pi",logTransform=FALSE,sigLine=NA, relative.positions = TRUE)
```

Visualizing pi across the genome for the Control treatment:

```{r}
dfm.pi <- m2.pi[[1]]
dfm.pi$chrom_alt <- factor(dfm.pi$chrom_alt, levels=c(0,1))

dfmsplit.pi <- split(dfm.pi, dfm.pi$chrom)
xbreaks <- sapply(dfmsplit.pi,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

m2.pi.total <- ggplot(dfm.pi, aes(x= index, y=marker, colour = as.factor(chrom_alt)))+
  geom_point(size=0.5, aes(alpha=0.33))+  
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm.pi$index)+10)) +
  guides(colour = FALSE,alpha=FALSE) +
  scale_y_continuous(expand = c(0,0),limits=c(0,max(dfm.pi$marker+0.001,na.rm=TRUE)))+
  labs(x = "Chromosome") +
  ggtitle("Control Treatment") +theme_classic()+
  labs(y=expression(pi)) +
  scale_color_manual(values=c("#56B4E9","dark blue")) 

m2.pi.total
```

```{r}
m2.pi <-m2.pi.total



m3.pi <-ggman::ggman(mydf.pi,chrom="CHR",bp="BIN_START",pvalue="STRESS_PI",snp="SNP.pi",logTransform=FALSE,sigLine=NA, ymax=1, relative.positions = TRUE)
```

Visualizing pi across the genome for the Stress treatment:

```{r}
dfm3.pi <- m3.pi[[1]]
dfm3.pi$chrom_alt <- factor(dfm3.pi$chrom_alt, levels=c(0,1))

dfmsplit <- split(dfm3.pi, dfm3.pi$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

m3.pi.total <- ggplot(dfm3.pi, aes(x= index, y=marker, colour = as.factor(chrom_alt)))+
  geom_point(size=0.5, aes(alpha=0.33))+  
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm3.pi$index)+10)) +
  guides(colour = FALSE,alpha=FALSE) +
  scale_y_continuous(expand = c(0,0),limits=c(0,max(dfm3.pi$marker+0.001,na.rm=TRUE)))+
  #scale_alpha_continuous(range=c(0.1,0.95))+
  labs(x = "Chromosome") +
  ggtitle("Stress Treatment") +theme_classic()+
  labs(y=expression(pi)) + 
  scale_color_manual(values=c("orange", "#D55E00"))

m3.pi.total
```

```{r}
m3.pi <-m3.pi.total


md4.pi <-ggman::ggman(mydf.pi,chrom="CHR",bp="BIN_START",pvalue="DIFF",snp="SNP.pi",logTransform=FALSE,sigLine=NA, relative.positions = TRUE)
```

Visualizing the difference in Pi between the Control and Stress treatments

```{r}
dfm4.pi <- md4.pi[[1]]
dfm4.pi$chrom_alt <- factor(dfm4.pi$chrom_alt, levels=c(0,1))

dfmsplit <- split(dfm4.pi, dfm4.pi$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

md4.pi <- ggplot(dfm4.pi, aes(x= index, y=marker, colour = as.factor(chrom_alt),size=(0.1+ abs(marker))^2))+
  geom_point(alpha=0.85)+  
  scale_size_continuous(range=c(0.01,1.5))+
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm4.pi$index)+10)) +
  guides(colour = FALSE,alpha=FALSE,size=FALSE) +
  scale_y_continuous(expand = c(0,0),limits=c(-0.001,0.001), labels = scales::label_number(accuracy = 0.0001, scale = 1, decimal.mark = ".", big.mark = ""))+
  #scale_alpha_continuous(range=c(0.1,0.95))+
  labs(x = "Chromosome") +
  ggtitle("Control - Stress") +theme_classic()+
  labs(y=expression("Difference in " ~ pi)) + 
  scale_color_manual(values=c("#009E73", "#00543d"))

md4.pi
```

```{r}
png(filename=paste(prefix.I,"Figure.pi_comparison.png", sep=""), type="cairo",units="px", width=5600, height=3000, res=300, bg="transparent")

layout <- "
AAADD
BBBDD
CCCEE
CCCEE
"
m2.pi.total+m3.pi.total+md4.pi+bd.pi+bc.pi +patchwork::plot_layout(design=layout, guides = "collect") + patchwork::plot_annotation(tag_levels = 'A')
```

Link to view plot: https://github.com/amyzyck/RISG_LarvJuvCADO/blob/main/Output/Figures/Figure.pi_comparison.png


### Tajima's D

In PATH: `/home/azyck/RISG_CADO/juv_ddocent/Juv_Selection_Analysis`

```{bash}
mkdir tajima
```

```{bash}
cd tajima
```

```{bash}
ln -s ../nucleotide_diversity/Control*.recode.vcf .
ln -s ../nucleotide_diversity/Stress*.recode.vcf .
```

```{bash}
conda create -n vk python=3.8 numpy scipy
conda activate vk
pip install VCF-kit
```

Calculating Tajima's D for each replicate. The first number is window-size and the second number is step-size

```{bash}
vk tajima 10000 10000 Control1.recode.vcf > Control1.tajima
vk tajima 10000 10000 Control2.recode.vcf > Control2.tajima
vk tajima 10000 10000 Control3.recode.vcf > Control3.tajima

vk tajima 10000 10000 Stress1.recode.vcf > Stress1.tajima
vk tajima 10000 10000 Stress2.recode.vcf > Stress2.tajima
vk tajima 10000 10000 Stress3.recode.vcf > Stress3.tajima
```

The following steps are identical to those above for organizing files and calculating mean windowed values

```{bash}
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_SITES\tN_SNPs\tTAJD\tPOP\tGROUP")  <(mawk '{if (NR>1)print $0 "\tCON1\tCONTROL"}' Control1.tajima)   <(mawk '{if (NR>1)print $0 "\tCON2\tCONTROL"}' Control2.tajima) <(mawk '{if (NR>1)print $0 "\tCON3\tCONTROL"}' Control3.tajima) > Control.windowed.tajima

cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_SITES\tN_SNPs\tTAJD\tPOP\tGROUP") <(mawk '{if (NR>1)print $0 "\tHI1\tSTRESS"}' Stress1.tajima)   <(mawk '{if (NR>1)print $0 "\tHI2\tSTRESS"}' Stress2.tajima) <(mawk '{if (NR>1)print $0 "\tHI3\tSTRESS"}' Stress3.tajima) > Stress.windowed.tajima

cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_SITES\tN_SNPs\tTAJD\tPOP\tGROUP")  <(mawk '{if (NR>1)print $0 "\tCON1\tCONTROL"}' Control1.tajima)   <(mawk '{if (NR>1)print $0 "\tCON2\tCONTROL"}' Control2.tajima) <(mawk '{if (NR>1)print $0 "\tCON3\tCONTROL"}' Control3.tajima) <(mawk '{if (NR>1)print $0 "\tHI1\tSTRESS"}' Stress1.tajima)   <(mawk '{if (NR>1)print $0 "\tHI2\tSTRESS"}' Stress2.tajima) <(mawk '{if (NR>1)print $0 "\tHI3\tSTRESS"}' Stress3.tajima) > all.windowed.tajima
```

```{bash}
bedtools merge -i <(mawk '!/CHR/' Control.windowed.tajima | sort -k1,1 -k2,2n ) -c 4,5,6,8 -d -1 -o mean,mean,mean,first > mean.Control.windowed.tajima

bedtools merge -i <(mawk '!/CHR/' Stress.windowed.tajima | sort -k1,1 -k2,2n ) -c 4,5,6,8 -d -1 -o mean,mean,mean,first > mean.Stress.windowed.tajima
```

```{bash}
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_SITES\tN_SNPs\tTAJD\tGROUP") mean.Control.windowed.tajima mean.Stress.windowed.tajima > mean.combined.tajima

cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_SITES\tN_SNPs\tTAJD\tGROUP") mean.Control.windowed.tajima > mean.Control.windowed.tajima.txt

cat <(echo -e "CHROM\tBIN_START\tBIN_END\tN_SITES\tN_SNPs\tTAJD\tGROUP") mean.Stress.windowed.tajima > mean.Stress.windowed.tajima.txt
```

```{r}
prefix.J <- "../tajima/"
```

```{r}
tajd.all.dataframe<-read.table(paste(prefix.J,"all.windowed.tajima", sep=""), sep="\t", header=T)
summary(tajd.all.dataframe)
```

```{r}
tajd.all.dataframe$GROUP <- factor(tajd.all.dataframe$GROUP, levels=c("CONTROL", "STRESS"))
tajd.all.dataframe$POP <- gsub("^HI", "STR", tajd.all.dataframe$POP)
tajd.all.dataframe$POP <- factor(
  tajd.all.dataframe$POP,
  levels = c("CON1", "CON2", "CON3", "STR1", "STR2", "STR3")
)


tajd.all.mean.dataframe<-read.table(paste(prefix.J,"mean.combined.tajima", sep=""), sep="\t", header=T)
tajd.all.mean.dataframe$GROUP <- factor(tajd.all.mean.dataframe$GROUP, levels=c("CONTROL", "STRESS"))
summary(tajd.all.mean.dataframe)
```

Plotting

```{r}
library(ggtext)
```

```{r}
bd.taj <-ggplot(tajd.all.dataframe, aes(x=TAJD,y = POP))+
  #geom_violin(aes(color=GROUP,fill= GROUP)) +
  #stat_summary(fun=mean, geom="point", shape=23, size=4)+
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=0.005)+
  geom_boxplot(aes(fill=GROUP), width=0.75,outlier.shape = 23, outlier.color = "black")+
  stat_summary(fun=mean, geom="point", shape=23, size=2)+
  scale_fill_manual(values=c("#56B4E9", "orange"))+
  scale_color_manual(values=c("#56B4E9", "orange"))+
  scale_y_discrete(limits = rev(levels(tajd.all.dataframe$POP)))+
  #ylab("Population/Line")+
  ggtitle("Control vs. Stress Tajima's *D*")+
  theme_classic() +
  labs(x="Tajima's *D*") +theme(axis.title.x = ggtext::element_markdown(),  axis.title.y = element_blank(), plot.title = ggtext::element_markdown()) +
  theme(legend.position = "none")

bd.taj <- bd.taj + guides(color = "none", fill ="none")
bd.taj
```

```{r}
bc.taj <-ggplot(tajd.all.mean.dataframe, aes(x=TAJD,y = GROUP))+
  geom_violin(aes(color=GROUP,fill= GROUP)) +
  geom_boxplot(aes(fill=GROUP), width=0.1,outlier.shape = 23, outlier.color = "black")+
  stat_summary(fun=mean, geom="point", shape=23, size=2)+
  scale_fill_manual(values=c("#56B4E9", "orange"))+
  scale_color_manual(values=c("#56B4E9", "orange"))+
  scale_y_discrete(limits = rev(levels(tajd.all.dataframe$GROUP)))+
  #ylab("Origin")+
  ggtitle("Mean Control vs. Stress Tajima's *D*")+
  theme_classic() +
  labs(x="Tajima's *D*") +
  theme(legend.position = c(0.1, 0.5), axis.title.y = element_blank(),axis.text.y=element_blank(), plot.title = ggtext::element_markdown(),axis.title.x = ggtext::element_markdown())

bc.taj
```

t-tests

```{r}
tajd.all.dataframe %>% t_test(formula = TAJD ~ GROUP, alternative = "two.sided",order = c("CONTROL", "STRESS"))
```

```{r}
tajd.all.dataframe %>% t_test(formula = TAJD ~ GROUP, alternative = "greater", order = c("CONTROL", "STRESS"))
```

Calculating summary statistics

```{r}
group_by(tajd.all.dataframe, GROUP) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(TAJD, na.rm = TRUE),
    sd = sd(TAJD, na.rm = TRUE),
    se=std.error(TAJD, na.rm = TRUE) )
```

```{r}
group_by(tajd.all.dataframe, POP) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(TAJD, na.rm = TRUE),
    sd = sd(TAJD, na.rm = TRUE),
    se=std.error(TAJD, na.rm = TRUE) )
```

**Plotting Tajima's D across the genome**

```{r}
filename <- paste(prefix.J,"mean.Control.windowed.tajima.txt", sep="")
ctajd <- read.table(filename, header = TRUE)
filename <- paste(prefix.J,"mean.Stress.windowed.tajima.txt", sep="")
stajd <- read.table(filename, header = TRUE)

ctajd <- ctajd %>% dplyr::rename(CONTROL_TAJD= TAJD)
stajd <- stajd %>% dplyr::rename(STRESS_TAJD= TAJD)
c.s.tajd <- plyr::join(ctajd,stajd, by =c("CHROM", "BIN_START"),type="full")


c.s.tajd$DIFF <- c.s.tajd$CONTROL_TAJD - c.s.tajd$STRESS_TAJD
```

```{r}
SNP<-c(1:(nrow(c.s.tajd)))
mydf<-data.frame(SNP,c.s.tajd)
mydf$CHR <- mydf$CHROM
mydf %>% 
  mutate(CHR = str_replace(CHR, "NC_035780.1", "1")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035781.1", "2")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035782.1", "3")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035783.1", "4")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035784.1", "5")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035785.1", "6")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035786.1", "7")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035787.1", "8")) %>%
  mutate(CHR = str_replace(CHR, "NC_035788.1", "9")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035789.1", "10"))  -> mydf
mydf$CHR <- as.numeric(mydf$CHR)  

m2 <-ggman::ggman(mydf,chrom="CHR",bp="BIN_START",pvalue="CONTROL_TAJD",snp="SNP",logTransform=FALSE,sigLine=NA, relative.positions = TRUE)
```

```{r}
dfm <- m2[[1]]
dfm$chrom_alt <- factor(dfm$chrom_alt, levels=c(0,1))

dfmsplit <- split(dfm, dfm$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

m2.total <- ggplot(dfm, aes(x= index, y=marker, colour = as.factor(chrom_alt)))+
  geom_point(size=0.5, aes(alpha=0.33))+  
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm$index)+10)) +
  guides(colour = "none",alpha="none") +
  scale_y_continuous(expand = c(0,0),limits=c(min(dfm$marker-0.001,na.rm=TRUE),max(dfm$marker+0.001,na.rm=TRUE)))+
  labs(x = "Chromosome") +
  ggtitle("Control Treatment") +theme_classic()+
  labs(y="Tajima's *D*") +theme(axis.title.y = ggtext::element_markdown())+ 
  scale_color_manual(values=c("#56B4E9","dark blue")) 
m2.taj <-m2.total

m2.taj
```

```{r}
m3 <-ggman::ggman(mydf,chrom="CHR",bp="BIN_START",pvalue="STRESS_TAJD",snp="SNP",logTransform=FALSE,sigLine=NA, ymax=1, relative.positions = TRUE)
```

```{r}
dfm <- m3[[1]]
dfm$chrom_alt <- factor(dfm$chrom_alt, levels=c(0,1))

dfmsplit <- split(dfm, dfm$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

m3.total <- ggplot(dfm, aes(x= index, y=marker, colour = as.factor(chrom_alt)))+
  geom_point(size=0.5, aes(alpha=0.33))+  
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm$index)+10)) +
  guides(colour = "none",alpha="none") +
  scale_y_continuous(expand = c(0,0),limits=c(min(dfm$marker-0.001,na.rm=TRUE),max(dfm$marker+0.001,na.rm=TRUE)))+
  #scale_alpha_continuous(range=c(0.1,0.95))+
  labs(x = "Chromosome") +
  ggtitle("Stress Treatment") +theme_classic()+
  labs(y="Tajima's *D*") +theme(axis.title.y = ggtext::element_markdown())+ 
  scale_color_manual(values=c("orange", "#D55E00"))

m3.taj <-m3.total

m3.taj
```

```{r}
md4 <-ggman::ggman(mydf,chrom="CHR",bp="BIN_START",pvalue="DIFF",snp="SNP",logTransform=FALSE,sigLine=NA, relative.positions = TRUE)
```

```{r}
y_range <- range(dfm$marker, na.rm = TRUE)
y_buffer <- 0.1 * diff(y_range)  # 10% buffer

md4.taj <- ggplot(dfm, aes(x= index, y=marker, colour = as.factor(chrom_alt), size=(0.1 + abs(marker))^2)) +
  geom_point(alpha=0.85) +
  scale_size_continuous(range = c(0.01, 1.5)) +
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks), expand = c(0, 0), limits = c(0, max(dfm$index) + 10)) +
  guides(colour = "none", alpha = "none", size = "none") +
  scale_y_continuous(expand = c(0, 0), limits = c(y_range[1] - y_buffer, y_range[2] + y_buffer)) +
  labs(x = "Chromosome") +
  ggtitle("Control - Stress") + theme_classic() +
  labs(y = "Difference in Tajima's *D*") + theme(axis.title.y = ggtext::element_markdown()) +
  scale_color_manual(values = c("#009E73", "#00543d"))
md4.taj
```



```{r}
png(filename=paste(prefix.J,"Figure.TajimasD.Comparison.Con.Str.png", sep=""), type="cairo",units="px", width=5600, height=3000, res=300, bg="transparent")
layout <- "
AAADD
BBBDD
CCCEE
CCCEE
"
#######(m2/m3/md4 | bd/bc) +plot_layout(design=layout)
m2.taj+m3.taj+md4.taj+bd.taj+bc.taj+patchwork::plot_layout(design=layout, guides = "collect") + patchwork::plot_annotation(tag_levels = 'A')
```


### Heterozygosity

```{bash}
conda activate vk
conda install bioconda::vcflib
```

Back in PATH: `/home/azyck/RISG_CADO/juv_ddocent/Juv_Selection_Analysis`

```{bash}
mkdir heterozygosity
cd heterozygosity
```

```{bash}
ln -s ../nucleotide_diversity/Control*.recode.vcf .
ln -s ../nucleotide_diversity/Stress*.recode.vcf .
```

Calculating heterozygosity using popStats

```{bash}
popStats -y GT -t 0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19 --file Control1.recode.vcf  > Control1.popstats
popStats -y GT -t 0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19 --file Control2.recode.vcf  > Control2.popstats
popStats -y GT -t 0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19 --file Control3.recode.vcf  > Control3.popstats

popStats -y GT -t 0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19 --file Stress1.recode.vcf  > Stress1.popstats
popStats -y GT -t 0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19 --file Stress2.recode.vcf  > Stress2.popstats
popStats -y GT -t 0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19 --file Stress3.recode.vcf  > Stress3.popstats
```

Format popstats to BED

```{bash,eval=FALSE}
bedtools makewindows -g ../reference.fasta.fai -w 10000 | mawk '!/NC_007175.2/' > 10kb.bed
AWK1='{print $1 "\t" $2-1 "\t" $2 "\t" $3 "\t" $4}'
ls ./*.popstats | sed 's/.popstats//g' | sed 's:\./::g' | parallel "cut -f1,2,4,5 ./{}.popstats | mawk '$AWK1' > ./{}.het.bed"
```

Aggregate per 10kb

```{bash,eval=FALSE}
ls ./*.het.bed | sed 's/.het.bed//g'  |sed 's:\./::g' | parallel "bedtools intersect -b ./{}.het.bed -a ./10kb.bed -wa -wb | bedtools merge -d -1 -c 7,8 -o mean > ./{}.het.per10kb.bed"
```

The following steps are the same as I did for nucleotide diversity and Tajima's D

```{bash} 
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tEXP_HET\tOBS_HET\tPOP\tGROUP") <(mawk '{if (NR>1)print $0 "\tCON1\tCONTROL"}' Control1.het.per10kb.bed)   <(mawk '{if (NR>1)print $0 "\tCON2\tCONTROL"}' Control2.het.per10kb.bed) <(mawk '{if (NR>1)print $0 "\tCON3\tCONTROL"}' Control3.het.per10kb.bed) > Control.windowed.het

cat <(echo -e "CHROM\tBIN_START\tBIN_END\tEXP_HET\tOBS_HET\tPOP\tGROUP")  <(mawk '{if (NR>1)print $0 "\tHI1\tSTRESS"}' Stress1.het.per10kb.bed)   <(mawk '{if (NR>1)print $0 "\tHI2\tSTRESS"}' Stress2.het.per10kb.bed) <(mawk '{if (NR>1)print $0 "\tHI3\tSTRESS"}' Stress3.het.per10kb.bed) > Stress.windowed.het

cat <(echo -e "CHROM\tBIN_START\tBIN_END\tEXP_HET\tOBS_HET\tPOP\tGROUP")   <(mawk '{if (NR>1)print $0 "\tCON1\tCONTROL"}' Control1.het.per10kb.bed)  <(mawk '{if (NR>1)print $0 "\tCON2\tCONTROL"}' Control2.het.per10kb.bed) <(mawk '{if (NR>1)print $0 "\tCON3\tCONTROL"}' Control3.het.per10kb.bed) <(mawk '{if (NR>1)print $0 "\tHI1\tSTRESS"}' Stress1.het.per10kb.bed)   <(mawk '{if (NR>1)print $0 "\tHI2\tSTRESS"}' Stress2.het.per10kb.bed) <(mawk '{if (NR>1)print $0 "\tHI3\tSTRESS"}' Stress3.het.per10kb.bed) > all.windowed.het
```

```{bash}
bedtools merge -i <(mawk '!/CHR/' Control.windowed.het| sort -k1,1 -k2,2n ) -c 4,5,7 -d -1 -o mean,mean,first > mean.Control.windowed.het
bedtools merge -i <(mawk '!/CHR/' Stress.windowed.het | sort -k1,1 -k2,2n ) -c 4,5,7 -d -1 -o mean,mean,first > mean.Stress.windowed.het
```

```{bash}
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tEXP_HET\tOBS_HET\tGROUP") mean.Control.windowed.het mean.Stress.windowed.het > mean.combined.windowed.het
```

```{bash}
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tEXP_HET\tOBS_HET\tGROUP") mean.Control.windowed.het > mean.Control.windowed.het.txt
cat <(echo -e "CHROM\tBIN_START\tBIN_END\tEXP_HET\tOBS_HET\tGROUP") mean.Stress.windowed.het > mean.Stress.windowed.het.txt
```

```{r}
prefix.H <- "../heterozygosity/"
```

```{r}
filename <- paste(prefix.H,"all.windowed.het", sep="")
het.all.dataframe<-read.table(filename, sep="\t", header=T)
summary(het.all.dataframe)
```


```{r}
het.all.dataframe$GROUP <- factor(het.all.dataframe$GROUP, levels=c("CONTROL", "STRESS"))
het.all.dataframe$POP <- gsub("^HI", "STR", het.all.dataframe$POP)
het.all.dataframe$POP <- factor(
  het.all.dataframe$POP,
  levels = c("CON1", "CON2", "CON3", "STR1", "STR2", "STR3")
)

filename <- paste(prefix.H,"mean.combined.windowed.het", sep="")
het.all.mean.dataframe<-read.table(filename, sep="\t", header=T)
het.all.mean.dataframe$GROUP <- factor(het.all.mean.dataframe$GROUP, levels=c("CONTROL", "STRESS"))
summary(het.all.mean.dataframe)
```


Plotting

```{r}
bd.het <-ggplot(het.all.dataframe, aes(x=OBS_HET,y = POP))+
  #geom_violin(aes(color=GROUP,fill= GROUP)) +
  #stat_summary(fun=mean, geom="point", shape=23, size=4)+
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=0.005)+
  geom_boxplot(aes(fill=GROUP), width=0.75,outlier.shape = NA, outlier.color = "black")+
  stat_summary(fun=mean, geom="point", shape=23, size=2)+
  scale_fill_manual(values=c("#56B4E9", "orange"))+
  scale_color_manual(values=c("#56B4E9", "orange"))+
  scale_y_discrete(limits = rev(levels(het.all.dataframe$POP)))+
  #ylab("Population/Line")+
  ggtitle("Control vs. Stress Heterozygosity")+
  coord_cartesian(xlim=c(0,0.5))+
  theme_classic() +
  labs(x="Observed Heterozygosity") +theme(axis.title.x = ggtext::element_markdown(),  axis.title.y = element_blank(), plot.title = ggtext::element_markdown()) +
  theme(legend.position = "none")


bd.het <- bd.het + guides(color = "none", fill ="none")
bd.het
```

```{r}
bc.het <-ggplot(het.all.mean.dataframe, aes(x=OBS_HET,y = GROUP))+
  geom_violin(aes(color=GROUP,fill= GROUP),trim=TRUE) +
  geom_boxplot(aes(fill=GROUP), width=0.1,outlier.shape = NA, outlier.color = "black")+
  stat_summary(fun=mean, geom="point", shape=23, size=2)+
  scale_fill_manual(values=c("#56B4E9", "orange"))+
  scale_color_manual(values=c("#56B4E9", "orange"))+
  scale_y_discrete(limits = rev(levels(het.all.dataframe$GROUP)))+
  #ylab("Origin")+
  ggtitle("Mean Control vs. Stress Heterozygosity")+
  theme_classic() +
  scale_x_continuous(breaks=c(0,0.1,0.2,0.3,0.4,0.5), labels=c("0.0","0.1","0.2","0.3","0.4","0.5*"))+
  coord_cartesian(xlim=c(0,0.5))+
  labs(x="Observed Heterozygosity") +
  theme(legend.position = c(0.1, 0.5), axis.title.y = element_blank(),axis.text.y=element_blank(), plot.title = ggtext::element_markdown(),axis.title.x = ggtext::element_markdown())
bc.het
```

t-tests

```{r}
het.all.mean.dataframe %>% t_test(formula = OBS_HET ~ GROUP, alternative = "two.sided",order = c("CONTROL", "STRESS"))
```

```{r}
het.all.mean.dataframe %>% t_test(formula = OBS_HET ~ GROUP, alternative = "greater", order = c("CONTROL", "STRESS"))
```

Calculating summary statistics

```{r}
group_by(het.all.mean.dataframe, GROUP) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(OBS_HET, na.rm = TRUE),
    sd = sd(OBS_HET, na.rm = TRUE),
    se=std.error(OBS_HET, na.rm = TRUE) )
```

```{r}
group_by(het.all.dataframe, POP) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(OBS_HET, na.rm = TRUE),
    sd = sd(OBS_HET, na.rm = TRUE),
    se=std.error(OBS_HET, na.rm = TRUE) )
```

**Plotting heterozygosity across the genome**

```{r}
filename <- paste(prefix.H,"mean.Control.windowed.het.txt", sep="")
chet <- read.table(filename, header = TRUE)
filename <- paste(prefix.H,"mean.Stress.windowed.het.txt", sep="")
shet <- read.table(filename, header = TRUE)

chet <- chet %>% dplyr::rename(CONTROL_OBS_HET= OBS_HET)
shet <- shet %>% dplyr::rename(STRESS_OBS_HET= OBS_HET)
c.s.het <- plyr::join(chet,shet, by =c("CHROM", "BIN_START"),type="full")


c.s.het$DIFF <- c.s.het$CONTROL_OBS_HET - c.s.het$STRESS_OBS_HET
```

```{r}
SNP<-c(1:(nrow(c.s.het)))
mydf<-data.frame(SNP,c.s.het)
mydf$CHR <- mydf$CHROM
mydf %>% 
  mutate(CHR = str_replace(CHR, "NC_035780.1", "1")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035781.1", "2")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035782.1", "3")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035783.1", "4")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035784.1", "5")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035785.1", "6")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035786.1", "7")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035787.1", "8")) %>%
  mutate(CHR = str_replace(CHR, "NC_035788.1", "9")) %>% 
  mutate(CHR = str_replace(CHR, "NC_035789.1", "10"))  -> mydf
mydf$CHR <- as.numeric(mydf$CHR)  

m2.het <-ggman::ggman(mydf,chrom="CHR",bp="BIN_START",pvalue="CONTROL_OBS_HET",snp="SNP",logTransform=FALSE,sigLine=NA, relative.positions = TRUE)
```

```{r}
dfm <- m2.het[[1]]
dfm$chrom_alt <- factor(dfm$chrom_alt, levels=c(0,1))

dfmsplit <- split(dfm, dfm$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

m2.total <- ggplot(dfm, aes(x= index, y=marker, colour = as.factor(chrom_alt)))+
  geom_point(size=0.5, aes(alpha=0.33))+  
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm$index)+10)) +
  guides(colour = "none",alpha="none") +
  scale_y_continuous(expand = c(0,0),limits=c(min(dfm$marker-0.001,na.rm=TRUE),0.9))+
  labs(x = "Chromosome") +
  ggtitle("Control Treatment") +theme_classic()+
  labs(y="Observed Heterozygosity") +theme(axis.title.y = ggtext::element_markdown())+ 
  scale_color_manual(values=c("#56B4E9","dark blue")) 
m2.het <- m2.total

m2.het
```

```{r}
m3.het <-ggman::ggman(mydf,chrom="CHR",bp="BIN_START",pvalue="STRESS_OBS_HET",snp="SNP",logTransform=FALSE,sigLine=NA, ymax=1, relative.positions = TRUE)
```

```{r}
dfm <- m3.het[[1]]
dfm$chrom_alt <- factor(dfm$chrom_alt, levels=c(0,1))

dfmsplit <- split(dfm, dfm$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

m3.total <- ggplot(dfm, aes(x= index, y=marker, colour = as.factor(chrom_alt)))+
  geom_point(size=0.5, aes(alpha=0.33))+  
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm$index)+10)) +
  guides(colour = "none",alpha="none") +
  scale_y_continuous(expand = c(0,0),limits=c(min(dfm$marker-0.001,na.rm=TRUE),0.9))+
  #scale_alpha_continuous(range=c(0.1,0.95))+
  labs(x = "Chromosome") +
  ggtitle("Stress Treatment") +theme_classic()+
  labs(y="Observed Heterozygosity") +theme(axis.title.y = ggtext::element_markdown())+ 
  scale_color_manual(values=c("orange", "#D55E00"))

m3.het <- m3.total

m3.het
```

```{r}
md4.het <-ggman::ggman(mydf,chrom="CHR",bp="BIN_START",pvalue="DIFF",snp="SNP",logTransform=FALSE,sigLine=NA, relative.positions = TRUE)
```

```{r}
dfm <- md4.het[[1]]
dfm$chrom_alt <- factor(dfm$chrom_alt, levels=c(0,1))

dfmsplit <- split(dfm, dfm$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

md4.het <- ggplot(dfm, aes(x= index, y=marker, colour = as.factor(chrom_alt),size=(0.1+ abs(marker))^2))+
  geom_point(alpha=0.85)+  
  scale_size_continuous(range=c(0.01,1.5))+
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm$index)+10)) +
  guides(colour = "none",alpha="none",size="none") +
  scale_y_continuous(expand = c(0,0),limits=c(-0.4,0.4))+
  #scale_alpha_continuous(range=c(0.1,0.95))+
  labs(x = "Chromosome") +
  ggtitle("Control - Stress") +theme_classic()+
  labs(y="Difference in Observed Heterozygosity") +theme(axis.title.y = ggtext::element_markdown())+ 
  scale_color_manual(values=c("#009E73", "#00543d"))

md4.het
```

```{r}
png(filename=paste(prefix.H,"Figure.Heterozygosity.Comparison.Con.Str.png", sep=""), type="cairo",units="px", width=5600, height=3000, res=300, bg="transparent")
layout <- "
AAADD
BBBDD
CCCEE
CCCEE
"
m2.het+m3.het+md4.het+bd.het+bc.het + patchwork::plot_layout(design=layout, guides = "collect") + patchwork::plot_annotation(tag_levels = 'A')
```

### Putting all violin plots together

```{r}
library(patchwork)
```

```{r}
# Remove legends from all but one plot
bc.het <- bc.het + theme(legend.position = "none")
bc.taj <- bc.taj + theme(legend.position = "none")

# Stack, collect legend, and add subplot labels
combined_plot <- bc.pi / bc.het / bc.taj +
  plot_layout(guides = "collect") &
  theme(legend.position = "right")

combined_plot <- combined_plot + plot_annotation(tag_levels = 'A')

# Show or save
print(combined_plot)
ggsave("geneticdiversity_combined_comparisonplot.png", combined_plot, width=8, height=12)
```
